---
title: "Decision tree"
output: html_document
date: "2022-08-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(data.table)
library(nnet)
library(partykit)
library(randomForest)
library(tidyverse)
library(tree)
library(anytime)
library(dplyr)
library(lubridate)
library(ggthemes)
library(knitr)
library(ggplot2)
library(mice)
library(lattice)
library(reshape2)
#install.packages("DataExplorer") if the following package is not available
library(DataExplorer)
library(visdat)
library(ROSE)
library(e1071)
library(pROC)
library(dplyr)
library(rpart)
library(rpart.plot)
library(Metrics)
library(mlr)
library(ggplot2)
library(plotly)


set.seed(100) # For reproducibility
```

# Input

```{r}
df_train <- read_csv("outputs/df_train3.csv", show_col_types = FALSE)
df_val <- read_csv("outputs/df_val3.csv", show_col_types = FALSE)
df_test <- read_csv("outputs/df_test3.csv", show_col_types = FALSE)

df_train = as.data.table(df_train)
df_train$month = as.factor(df_train$month)
df_train$weekdays = as.factor(df_train$weekdays)
df_train$weekend = as.factor(df_train$weekend)
df_train$quarter = as.factor(df_train$quarter)
df_train$week = format(df_train$date, "%V")
df_train = as.data.frame(df_train)
df_train$week = as.numeric(df_train$week)
df_train$state = as.factor(df_train$state)
df_train


df_val = as.data.table(df_val)
df_val$month = as.factor(df_val$month)
df_val$weekdays = as.factor(df_val$weekdays)
df_val$weekend = as.factor(df_val$weekend)
df_val$quarter = as.factor(df_val$quarter)
df_val$week = format(df_val$date, "%V")
df_val = as.data.frame(df_val)
df_val$week = as.numeric(df_val$week)
df_val$state = as.factor(df_val$state)
df_val

df_test = as.data.table(df_test)
df_test$month = as.factor(df_test$month)
df_test$weekdays = as.factor(df_test$weekdays)
df_test$weekend = as.factor(df_test$weekend)
df_test$quarter = as.factor(df_test$quarter)
df_test$week = format(df_test$date, "%V")
df_test = as.data.frame(df_test)
df_test$week = as.numeric(df_test$week)
df_test$state = as.factor(df_test$state)


df_train$mday <- df_train$day 
df_train$wday <- df_train$weekdays
df_train$hr <- df_train$Hour
df_train$qrtr <- df_train$quarter
df_train$mnth <- df_train$month
df_train$wend <- df_train$weekend
df_train$rmean <- df_train$rolling_mean
df_train$rmed <- df_train$rolling_median
df_train$rskew <- df_train$rolling_skewness
df_train$rkurt <- df_train$rolling_kurtosis
df_train$rstd <- df_train$rolling_std
df_train$rslope <- df_train$rolling_slope
df_train$gas <- df_train$kwh
df_train$l1 <- df_train$lag_1
df_train$l2 <- df_train$lag_2
df_train$l3 <- df_train$lag_3
df_train$l4 <- df_train$lag_4
df_train$l5 <- df_train$lag_5

df_val$mday <- df_val$day 
df_val$wday <- df_val$weekdays
df_val$hr <- df_val$Hour
df_val$qrtr <- df_val$quarter
df_val$mnth <- df_val$month
df_val$wend <- df_val$weekend
df_val$rmean <- df_val$rolling_mean
df_val$rmed <- df_val$rolling_median
df_val$rskew <- df_val$rolling_skewness
df_val$rkurt <- df_val$rolling_kurtosis
df_val$rstd <- df_val$rolling_std
df_val$rslope <- df_val$rolling_slope
df_val$gas <- df_val$kwh
df_val$l1 <- df_val$lag_1
df_val$l2 <- df_val$lag_2
df_val$l3 <- df_val$lag_3
df_val$l4 <- df_val$lag_4
df_val$l5 <- df_val$lag_5

df_test$mday <- df_test$day 
df_test$wday <- df_test$weekdays
df_test$hr <- df_test$Hour
df_test$qrtr <- df_test$quarter
df_test$mnth <- df_test$month
df_test$wend <- df_test$weekend
df_test$rmean <- df_test$rolling_mean
df_test$rmed <- df_test$rolling_median
df_test$rskew <- df_test$rolling_skewness
df_test$rkurt <- df_test$rolling_kurtosis
df_test$rstd <- df_test$rolling_std
df_test$rslope <- df_test$rolling_slope
df_test$gas <- df_test$kwh
df_test$l1 <- df_test$lag_1
df_test$l2 <- df_test$lag_2
df_test$l3 <- df_test$lag_3
df_test$l4 <- df_test$lag_4
df_test$l5 <- df_test$lag_5
```

```{r}
train <- subset(df_train, select = c('hr','yday','mday','wday','wend','week','mnth','qrtr','t','gas',
                                   'l1','l2','l3','l4','l5','rmean','rmed','rstd','rslope','rskew','rkurt',
                                   'state'))
val <- subset(df_val, select = c('hr','yday','mday','wday','wend','week','mnth','qrtr','t','gas',
                                   'l1','l2','l3','l4','l5','rmean','rmed','rstd','rslope','rskew','rkurt',
                                   'state'))
test <- subset(df_test, select = c('hr','yday','mday','wday','wend','week','mnth','qrtr','t','gas',
                                   'l1','l2','l3','l4','l5','rmean','rmed','rstd','rslope','rskew','rkurt',
                                   'state'))
```

```{r}
#Separate X and Y (for validation)
X_train <- train %>% 
  select(!contains(c("state")))
X_val <- val %>% 
  select(!contains(c("state")))
X_test <- test %>% 
  select(!contains(c("state")))

y_train <- train %>% select('state')
y_val <- val %>% select('state')
y_test <- test %>% select('state')

X_train
y_train

X_val
y_val

X_test
y_test
```

# Tree0

```{r}
d.tree0 = rpart(state ~ ., 
               data=train, 
               method = 'class')
```

As we are not specifying *hyperparameters*, we are using [rpart's](https://www.rdocumentation.org/packages/rpart/versions/4.1.16/topics/rpart) default values:

-   Our tree can descend until 30 levels --- `maxdepth = 30` ;

-   The minimum number of examples in a node to perform a split is 20 --- `minsplit = 20` ;

-   The minimum number of examples in a terminal node is 7 --- `minbucket = 7`;

-   The split must increase the "performance" (although it's not that direct, we can consider "performance" a proxy for `cp`) of the tree by at least 0.01 --- `cp = 0.01` ;

```{r}
png(paste0("tree1.png"), width = 150, height = 75, units='mm', res = 700)

rpart.plot(d.tree0, cex=0.7)

```

```{r}
# Predict Values
predicted_values0 <- predict(d.tree0, X_val, type = 'class')
# Getting Accuracy
accuracy(y_val$state, predicted_values0)
```

```{r}
actual <- y_val$state
pred <- predicted_values0

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
plot.roc(y_val$state,
         as.numeric(predicted_values0), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

#### tree0 test

```{r}
# Predict Values
predicted_values0 <- predict(d.tree0, X_test, type = 'class')
# Getting Accuracy
accuracy(y_test$state, predicted_values0)
```

```{r}
actual <- y_test$state
pred <- predicted_values0

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
cm_d <- 
  as.data.frame(as.table(confusionMatrix(pred, actual, mode = "everything", positive="1")))

cm_d <- cm_d %>% rename(Actual = Reference)
cm_d$classes <- 2
cm_d$model <- "Decision Tree"

png(paste0("DT_cm1.png"), width = 125, height = 75, units='mm', res = 700)

ggplot(cm_d, aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient2(guide=FALSE,low='grey70',high='gray45',midpoint=10000) + 
  guides(fill = "none") + labs(title = "Unbalanced") + 
  geom_text(aes(label = Freq), color = "black") + theme_tufte() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot.roc(y_test$state,
         as.numeric(predicted_values0), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

# Tree0 - balanced

```{r}
over <- ovun.sample(state~., data = train, method = "over", N = 102282)$data
table(over$state)
```

```{r}
d.tree0.bal = rpart(state ~ ., 
               data=over, 
               method = 'class')
```

```{r}
rpart.plot(d.tree0.bal, cex=0.8)
```

```{r}
# Predict Values
predicted_values0.bal <- predict(d.tree0.bal, X_val, type = 'class')
# Getting Accuracy
accuracy(y_val$state, predicted_values0.bal)
```

```{r}
actual <- y_val$state
pred <- predicted_values0.bal

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
plot.roc(y_val$state,
         as.numeric(predicted_values0.bal), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

#### tree0 bal test

```{r}
# Predict Values
predicted_values0.bal <- predict(d.tree0.bal, X_test, type = 'class')
# Getting Accuracy
accuracy(y_test$state, predicted_values0.bal)
```

```{r}
actual <- y_test$state
pred <- predicted_values0.bal

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
cm_d <- 
  as.data.frame(as.table(confusionMatrix(pred, actual, mode = "everything", positive="1")))

cm_d <- cm_d %>% rename(Actual = Reference)
cm_d$classes <- 2
cm_d$model <- "Decision Tree"

png(paste0("DT_cm2.png"), width = 125, height = 75, units='mm', res = 700)

ggplot(cm_d, aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient2(guide=FALSE,low='grey70',high='gray45',midpoint=10000) + 
  guides(fill = "none") + labs(title = "Balanced") + 
  geom_text(aes(label = Freq), color = "black") + theme_tufte() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot.roc(y_test$state,
         as.numeric(predicted_values0.bal), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

# Tree1

maxdepth=29; cp=0.001; minsplit=3 : f1.test.mean=0.8836364

```{r}
d.tree.custom = rpart(state~ ., 
               data=train, 
               method = 'class',
               control = c(maxdepth = 5, minsplit = 3, cp = 0.001))
```

```{r}
rpart.plot(d.tree.custom, cex=0.55)
```

```{r}
# Predict Values
predicted_values.cust <- predict(d.tree.custom, X_val, type = 'class')
# Getting Accuracy
accuracy(y_val$state, predicted_values.cust)
```

```{r}
actual <- y_val$state
pred <- predicted_values.cust

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
plot.roc(y_val$state,
         as.numeric(predicted_values.cust), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

#### tree1 test

```{r}
# Predict Values
predicted_values.cust <- predict(d.tree.custom, X_test, type = 'class')
# Getting Accuracy
accuracy(y_test$state, predicted_values.cust)
```

```{r}
actual <- y_test$state
pred <- predicted_values.cust

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
cm_d <- 
  as.data.frame(as.table(confusionMatrix(pred, actual, mode = "everything", positive="1")))

cm_d <- cm_d %>% rename(Actual = Reference)
cm_d$classes <- 2
cm_d$model <- "Decision Tree"

png(paste0("DTt_cm1.png"), width = 125, height = 75, units='mm', res = 700)

ggplot(cm_d, aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient2(guide=FALSE,low='grey70',high='gray45',midpoint=10000) + 
  guides(fill = "none") + labs(title = "Unbalanced") + 
  geom_text(aes(label = Freq), color = "black") + theme_tufte() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot.roc(y_test$state,
         as.numeric(predicted_values.cust), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

# Tree1 - balanced

```{r}
over <- ovun.sample(state~., data = train, method = "over", N = 102282)$data
table(over$state)
```

```{r}
d.tree.custom.bal = rpart(state~ ., 
               data=over, 
               method = 'class',
               control = c(maxdepth = 5, minsplit = 3, cp = 0.001))
```

```{r}
png(paste0("tree2.png"), width = 220, height = 80, units='mm', res = 700)

rpart.plot(d.tree.custom.bal, cex=0.7)
```

```{r}
# Predict Values
predicted_values.cust.bal <- predict(d.tree.custom.bal, X_val, type = 'class')
# Getting Accuracy
accuracy(y_val$state, predicted_values.cust.bal)
```

```{r}
actual <- y_val$state
pred <- predicted_values.cust.bal

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
plot.roc(y_val$state,
         as.numeric(predicted_values.cust.bal), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

#### tree1 bal test

```{r}
# Predict Values
predicted_values.cust.bal <- predict(d.tree.custom.bal, X_test, type = 'class')
# Getting Accuracy
accuracy(y_test$state, predicted_values.cust.bal)
```

```{r}
actual <- y_test$state
pred <- predicted_values.cust.bal

confusionMatrix(pred, actual, mode = "everything", positive="1")
```

```{r}
cm_d <- 
  as.data.frame(as.table(confusionMatrix(pred, actual, mode = "everything", positive="1")))

cm_d <- cm_d %>% rename(Actual = Reference)
cm_d$classes <- 2
cm_d$model <- "Decision Tree"

png(paste0("DTt_cm2.png"), width = 125, height = 75, units='mm', res = 700)

ggplot(cm_d, aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient2(guide=FALSE,low='grey70',high='gray45',midpoint=10000) + 
  guides(fill = "none") + labs(title = "Balanced") + 
  geom_text(aes(label = Freq), color = "black") + theme_tufte() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot.roc(y_test$state,
         as.numeric(predicted_values.cust.bal), col = "blue",
         lwd = 3, print.auc = TRUE, print.auc.y = 0.3,
         main = "ROC-AUC of DT")
```

```{r}
# <- as.data.frame(d.tree.custom.bal$variable.importance)
a <- varImp(d.tree.custom.bal, scale = FALSE)
a
```
