---
title: "Binary Classification"
output: github_document
---

Import Libraries
```{r}
library(caret)
library(corrplot)
library(data.table)
library(ggplot2)
library(gridExtra)
library(hrbrthemes)
library(nnet)
library(partykit)
library(randomForest)
library(tidyverse)
library(tree)
library(xgboost)
library(viridis)

set.seed(100) # For reproducibility
```

Load datasets
```{r}
df_consumption_daily_2011_2012 <- read_csv(
  "../data/household_consumption_2011_2012_daily.csv", show_col_types = FALSE)
df_consumption_daily_2013 <- read_csv(
  "../data/household_consumption_2013_daily.csv", show_col_types = FALSE)
df_consumption_daily_2014 <- read_csv(
  "../data/household_consumption_2014_daily.csv", show_col_types = FALSE)
df_households <- read_csv("../data/household_information.csv", 
                          show_col_types = FALSE)
df_weather_daily <- read_csv("../data/weather_daily.csv", 
                             show_col_types = FALSE)
```

Daily consumption data
```{r}
df_consumption_daily <- rbind(df_consumption_daily_2011_2012, 
                              df_consumption_daily_2013, 
                              df_consumption_daily_2014)
# Select only household ID, date, and average daily consumption
df_consumption_daily <- df_consumption_daily %>% 
  select(LCLid, date, kwh_per_hh_avg)
head(df_consumption_daily)
```
Household information
```{r}
head(df_households)
```

Weather dataset
```{r}
# Only select date and tavg
df_weather_daily <- df_weather_daily %>% select(date, tavg)
head(df_weather_daily)
```
```{r}
classes <- c("Adversity", "Affluent")
classes_level <- c(0, 1)
```

Clean household information dataset
```{r}
df_households <- df_households %>% select(-Acorn, -file)
df_households <- df_households[df_households$Acorn_grouped %in% c("Affluent", 
                                                                  "Adversity"),]
df_households[df_households$stdorToU == "ToU", ]$stdorToU <- "1"
df_households[df_households$stdorToU == "Std", ]$stdorToU <- "0"
df_households$stdorToU = as.integer(df_households$stdorToU)
colnames(df_households)[2] <- "is_ToU"
```

Join the consumption dataset with weather dataset
```{r}
df_joined_daily <- left_join(df_consumption_daily, df_weather_daily, 
                             by = "date")
head(df_joined_daily)
```

Data preprocessing
```{r}
df_joined_daily$week <- week(df_joined_daily$date)
df_joined_daily$kwh_per_temperature_avg = 
  df_joined_daily$kwh_per_hh_avg / df_joined_daily$tavg
head(df_joined_daily)
```

```{r}
df_joined_daily_summary <- inner_join(df_joined_daily, 
                                      df_households, 
                                      by = "LCLid")
df_joined_daily_summary <- df_joined_daily_summary %>% 
  group_by(week, Acorn_grouped) %>% 
  summarise(kwh_per_hh_avg = mean(kwh_per_hh_avg, na.rm = TRUE), 
            kwh_per_temperature_avg = mean(kwh_per_temperature_avg, 
                                           na.rm = TRUE)) %>% 
  ungroup()
summary(df_joined_daily_summary)
```

Without temperature
```{r}
df_joined_without_temperature <- df_joined_daily %>% select(LCLid, week, 
                                                            kwh_per_hh_avg)
df_joined_without_temperature <- df_joined_without_temperature %>% 
  group_by(LCLid, week) %>% summarise(kwh_per_hh_avg = mean(kwh_per_hh_avg, 
                                                            na.rm = TRUE)) %>% 
  ungroup()
df_joined_without_temperature <- pivot_wider(
  df_joined_without_temperature, names_from = week, 
  names_glue = "week_{week}_avg_exclude_temperature", 
  values_from = kwh_per_hh_avg)
# Remove NAs
df_joined_without_temperature <- df_joined_without_temperature %>% drop_na()
```

With temperature
```{r}
df_joined_with_temperature <- df_joined_daily %>% 
  select(LCLid, week, kwh_per_temperature_avg)
df_joined_with_temperature <- df_joined_with_temperature %>% 
  group_by(LCLid, week) %>% 
  summarise(kwh_per_temperature_avg = mean(kwh_per_temperature_avg, 
                                           na.rm = TRUE)) %>% ungroup()
df_joined_with_temperature <- pivot_wider(
  df_joined_with_temperature, names_from = week, 
  names_glue = "week_{week}_avg_include_temperature", 
  values_from = kwh_per_temperature_avg)
# Remove NAs
df_joined_with_temperature <- df_joined_with_temperature %>% drop_na()
```

Label
```{r}
label_ref <- data.frame(levels = c("Adversity", 
                                   "Affluent"), 
                        label = c(0, 
                                  1
                                  ))
```

Joined datasets
```{r}
df_joined <- left_join(df_joined_with_temperature, 
                        df_joined_without_temperature, by = "LCLid")
df_joined <- inner_join(df_joined, df_households, by = "LCLid")
df_joined <- left_join(df_joined, label_ref, by = c("Acorn_grouped" = "levels"))
head(df_joined)
```

Create train and test datasets
```{r}
df_joined <- df_joined %>% 
  mutate(n = row_number()) %>% 
  select(n, everything()) 
df_train <- df_joined %>% 
  group_by(Acorn_grouped) %>% 
  sample_frac(.75) %>% 
  ungroup()
df_test <- anti_join(df_joined, df_train, by = "LCLid")
df_train <- df_train %>% select(-n)
df_test <- df_test %>% select(-n)
```

Separate df_train and df_test
```{r}
# Without temperature
levels <- c("Adversity", "Affluent")
# Train
df_train_without_temperature <- df_train %>% 
  select(contains(c("exclude", "Acorn")))
df_train_without_temperature$Acorn_grouped <- factor(
  df_train_without_temperature$Acorn_grouped, levels = levels
)

# Test
df_test_without_temperature <- df_test %>% 
  select(contains(c("exclude", "Acorn")))
df_test_without_temperature$Acorn_grouped <- factor(
  df_test_without_temperature$Acorn_grouped, levels = levels
)

# With temperature
# Train
df_train_with_temperature <- df_train %>% 
  select(contains(c("include", "Acorn")))
df_train_with_temperature$Acorn_grouped <- factor(
  df_train_with_temperature$Acorn_grouped, levels = levels
)

# Test
df_test_with_temperature <- df_test %>% 
  select(contains(c("include", "Acorn")))
df_test_with_temperature$Acorn_grouped <- factor(
  df_test_with_temperature$Acorn_grouped, levels = levels
)
```

Separate X and y
```{r}
# Without temperature
X_train_without_temperature <- df_train %>% 
  select(contains(c("exclude")))
X_test_without_temperature <- df_test %>% 
  select(contains(c("exclude")))

# With temperature
X_train_with_temperature <- df_train %>% select(contains(c("include")))
X_test_with_temperature <- df_test %>% select(contains(c("include")))

y_train <- df_train %>% select(label)
y_test <- df_test %>% select(label)
```

Logistic regression
```{r}
# Without temperature - Train
logistic_without_temperature <- multinom(Acorn_grouped ~ ., 
                                         data = df_train_without_temperature)
logistic_without_temperature_pred_train <- logistic_without_temperature %>% 
  predict(df_train_without_temperature)
mean(logistic_without_temperature_pred_train == 
       df_train_without_temperature$Acorn_grouped)
```

```{r}
# Without temperature - Test
logistic_without_temperature_pred_test <- logistic_without_temperature %>% 
  predict(df_test_without_temperature)
accuracy_logistic_without_temperature_pred_test <- 
  mean(logistic_without_temperature_pred_test == 
       df_test_without_temperature$Acorn_grouped)
accuracy_logistic_without_temperature_pred_test
```
```{r}
cm_logistic_without_temperature_test <- as.data.frame(as.table(confusionMatrix(
  data = factor(logistic_without_temperature_pred_test, levels = classes), 
  reference = factor(df_test$Acorn_grouped, levels = classes),
  )))
cm_logistic_without_temperature_test <- cm_logistic_without_temperature_test %>% 
  rename(Actual = Reference)
cm_logistic_without_temperature_test$classes <- 2
cm_logistic_without_temperature_test$model <- "Logistic Regression"
cm_logistic_without_temperature_test$with_or_without_temperature <- 
  "Without Temperature"
cm_logistic_without_temperature_test$accuracy <- 
  accuracy_logistic_without_temperature_pred_test

png(paste0("output/Confusion Matrix - ", 
           "Logistic Regression - 2 Classes - Without temperature.png"), 
    width = 125, height = 75, units='mm', res = 700)

ggplot(cm_logistic_without_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Accuracy: ", 
                      round(cm_logistic_without_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
confusionMatrix(
  data = factor(logistic_without_temperature_pred_test, levels = classes),
  reference = factor(df_test$Acorn_grouped, levels = classes),
  mode = "everything")
```

```{r}
# With temperature - Train
logistic_with_temperature <- multinom(Acorn_grouped ~ ., 
                                      data = df_train_with_temperature)
logistic_with_temperature_pred_train <- logistic_with_temperature %>% 
  predict(df_train_with_temperature)
mean(logistic_with_temperature_pred_train == 
       df_train_with_temperature$Acorn_grouped)
summary(logistic_with_temperature_pred_train)
```

```{r}
# With temperature - Test
logistic_with_temperature_pred_test <- logistic_with_temperature %>% 
  predict(df_test_with_temperature)
accuracy_logistic_with_temperature_pred_test <- mean(
  logistic_with_temperature_pred_test == df_test$Acorn_grouped)
accuracy_logistic_with_temperature_pred_test
```

```{r}
cm_logistic_with_temperature_test <- as.data.frame(as.table(confusionMatrix(
  data = factor(logistic_with_temperature_pred_test, levels = classes),
  reference = factor(df_test$Acorn_grouped, levels = classes),
  )))
cm_logistic_with_temperature_test <- cm_logistic_with_temperature_test %>% 
  rename(Actual = Reference)
cm_logistic_with_temperature_test$classes <- 2
cm_logistic_with_temperature_test$model <- "Logistic Regression"
cm_logistic_with_temperature_test$with_or_without_temperature <- 
  "With Temperature"
cm_logistic_with_temperature_test$accuracy <- 
  accuracy_logistic_with_temperature_pred_test

png(paste0("output/Confusion Matrix -", 
           "Logistic Regression - 2 Classes - With temperature.png"), 
    width = 125, height = 75, units='mm', res = 700)

ggplot(cm_logistic_with_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Accuracy: ", 
                      round(cm_logistic_with_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
confusionMatrix(
  data = factor(logistic_with_temperature_pred_test, levels = classes),
  reference = factor(df_test$Acorn_grouped, levels = classes),
  mode = "everything"
  )
```

```{r}
plot_logistic_without_temperature <- 
  ggplot(cm_logistic_without_temperature_test, 
         aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill= "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Without temperature \n Accuracy: ", 
                      round(cm_logistic_without_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))

plot_logistic_with_temperature <- ggplot(cm_logistic_with_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("With temperature \n Accuracy: ", 
                      round(cm_logistic_with_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))

png("output/Confusion Matrix - Logistic Regression - 2 Classes - All.png", 
    width = 175, height = 75, units='mm', res = 700)

grid.arrange(plot_logistic_without_temperature, 
             plot_logistic_with_temperature,
             ncol = 2) 
```

Decision Tree
```{r}
# # Without temperature - Train
# dtree_without_temperature <- ctree(Acorn_grouped ~ ., 
#                                   data = df_train_without_temperature)
# dtree_without_temperature_pred_train <- dtree_without_temperature %>% 
#   predict(df_train_without_temperature)
# mean(dtree_without_temperature_pred_train == 
#        df_train_without_temperature$Acorn_grouped)
# ```
# ```{r}
# # Without temperature - Test
# dtree_without_temperature_pred_test <- dtree_without_temperature %>% 
#   predict(df_test_without_temperature)
# accuracy_dtree_without_temperature_pred_test <- 
#   mean(dtree_without_temperature_pred_test == 
#        df_test$Acorn_grouped)
# accuracy_dtree_without_temperature_pred_test
```

```{r}
# cm_dtree_without_temperature_test <- as.data.frame(as.table(confusionMatrix(
#   factor(dtree_without_temperature_pred_test),
#   factor(df_test_without_temperature$Acorn_grouped))))
# cm_dtree_without_temperature_test <- cm_dtree_without_temperature_test %>% 
#   rename(Actual = Reference)
# cm_dtree_without_temperature_test$classes <- 2
# cm_dtree_without_temperature_test$model <- "Decision Tree"
# cm_dtree_without_temperature_test$with_or_without_temperature <- 
#   "Without Temperature"
# cm_dtree_without_temperature_test$accuracy <- 
#   accuracy_dtree_without_temperature_pred_test
# 
# ggplot(cm_dtree_without_temperature_test, 
#        aes(x = Actual, y = Prediction, fill = Freq)) + 
#   geom_tile() + 
#   coord_equal() +
#   scale_fill_gradient() + 
#   guides(fill=F) + 
#   geom_text(aes(label = Freq), color = "white") + 
#   labs(title = paste0("Accuracy: ", 
#                       round(cm_dtree_without_temperature_test$accuracy, 3))) + 
#   theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# # With temperature - Train
# dtree_with_temperature <- ctree(Acorn_grouped ~ ., 
#                                 data = df_train_with_temperature)
# dtree_with_temperature_pred_train <- dtree_with_temperature %>% 
#   predict(df_train_with_temperature)
# mean(dtree_with_temperature_pred_train == 
#        df_train_with_temperature$Acorn_grouped)
# ```
# ```{r}
# # With temperature - Test
# dtree_with_temperature <- ctree(Acorn_grouped ~ ., 
#                                 data = df_train_with_temperature)
# dtree_with_temperature_pred_test <- dtree_with_temperature %>% 
#   predict(df_test_with_temperature)
# accuracy_dtree_with_temperature_pred_test <- 
#   mean(dtree_with_temperature_pred_test == df_test$Acorn_grouped)
# accuracy_dtree_with_temperature_pred_test
```

```{r}
# cm_dtree_with_temperature_test <- as.data.frame(as.table(confusionMatrix(
#   factor(dtree_with_temperature_pred_test),
#   factor(df_test$Acorn_grouped))))
# cm_dtree_with_temperature_test <- cm_dtree_with_temperature_test %>% 
#   rename(Actual = Reference)
# cm_dtree_with_temperature_test$classes <- 2
# cm_dtree_with_temperature_test$model <- "Decision Tree"
# cm_dtree_with_temperature_test$with_or_without_temperature <- 
#   "With Temperature"
# cm_dtree_with_temperature_test$accuracy <- 
#   accuracy_dtree_with_temperature_pred_test
# 
# ggplot(cm_dtree_with_temperature_test, 
#        aes(x = Actual, y = Prediction, fill = Freq)) + 
#   geom_tile() + 
#   coord_equal() +
#   scale_fill_gradient() + 
#   guides(fill=F) + 
#   geom_text(aes(label = Freq), color = "white") + 
#   labs(title = paste0("Accuracy: ", 
#                       round(cm_dtree_with_temperature_test$accuracy, 3))) + 
#   theme(plot.title = element_text(hjust = 0.5))
```

Random Forest
```{r}
# Without temperature - Train
rf_without_temperature <- randomForest(Acorn_grouped ~ ., 
                                       data = df_train_without_temperature)
rf_without_temperature_pred_train <- rf_without_temperature %>% 
  predict(df_train_without_temperature)
mean(rf_without_temperature_pred_train == 
       df_train_without_temperature$Acorn_grouped)
```

```{r}
# Without temperature - Test
rf_without_temperature_pred_test <- rf_without_temperature %>% 
  predict(df_test_without_temperature)
accuracy_rf_without_temperature_pred_test <- 
  mean(rf_without_temperature_pred_test == df_test$Acorn_grouped)
accuracy_rf_without_temperature_pred_test
```

```{r}
cm_rf_without_temperature_test <- as.data.frame(as.table(confusionMatrix(
  data = factor(rf_without_temperature_pred_test, levels = classes),
  reference = factor(df_test_without_temperature$Acorn_grouped, 
                     levels = classes), 
  )))
cm_rf_without_temperature_test <- cm_rf_without_temperature_test %>% 
  rename(Actual = Reference)
cm_rf_without_temperature_test$classes <- 2
cm_rf_without_temperature_test$model <- "Random Forest"
cm_rf_without_temperature_test$with_or_without_temperature <- 
  "Without Temperature"
cm_rf_without_temperature_test$accuracy <- 
  accuracy_rf_without_temperature_pred_test

png(paste0("output/Confusion Matrix - ", 
           "Random Forest - 2 Classes - Without temperature.png"),
    width = 125, height = 75, units='mm', res = 700)

ggplot(cm_rf_without_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Accuracy: ", 
                      round(cm_rf_without_temperature_test$accuracy, 3))) + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
confusionMatrix(
  data = factor(rf_without_temperature_pred_test, levels = classes),
  reference = factor(df_test$Acorn_grouped, levels = classes), 
  mode = "everything")
```

```{r}
# With temperature - Train
rf_with_temperature <- randomForest(Acorn_grouped ~ ., 
                                     data = df_train_with_temperature)
rf_with_temperature_pred_train <- rf_with_temperature %>% 
  predict(df_train_with_temperature)
mean(rf_with_temperature_pred_train == 
       df_train_with_temperature$Acorn_grouped)
```


```{r}
# With temperature - Test
rf_with_temperature_pred_test <- rf_with_temperature %>% 
  predict(df_test_with_temperature)
accuracy_rf_with_temperature_pred_test <- mean(
  rf_with_temperature_pred_test == df_test$Acorn_grouped)
accuracy_rf_with_temperature_pred_test
```

```{r}
# Plot confusion matrix
cm_rf_with_temperature_test <- as.data.frame(as.table(confusionMatrix(
  data = factor(rf_with_temperature_pred_test, levels = classes),
  reference = factor(df_test_with_temperature$Acorn_grouped, levels = classes)
  )))
cm_rf_with_temperature_test <- cm_rf_with_temperature_test %>% 
  rename(Actual = Reference)
cm_rf_with_temperature_test$classes <- 2
cm_rf_with_temperature_test$model <- "Random Forest"
cm_rf_with_temperature_test$with_or_without_temperature <- 
  "With Temperature"
cm_rf_with_temperature_test$accuracy <- 
  accuracy_rf_with_temperature_pred_test

png(paste0("output/Confusion Matrix - ", 
           "Random Forest - 2 Classes - With temperature.png"), 
    width = 125, height = 75, units='mm', res = 700)

ggplot(cm_rf_with_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Accuracy: ", 
                      round(cm_rf_with_temperature_test$accuracy, 3))) + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
confusionMatrix(
  data = factor(rf_with_temperature_pred_test, levels = classes),
  reference = factor(df_test$Acorn_grouped, levels = classes), 
  mode = "everything")
```

```{r}
plot_rm_without_temperature <- 
  ggplot(cm_rf_without_temperature_test, 
         aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Without temperature \n Accuracy: ", 
                      round(cm_rf_without_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))

plot_rm_with_temperature <- ggplot(cm_rf_with_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("With temperature \n Accuracy: ", 
                      round(cm_rf_with_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))

png("output/Confusion Matrix - Random Forest - 2 Classes - All.png", 
    width = 175, height = 75, units='mm', res = 700)

grid.arrange(plot_rm_without_temperature, 
             plot_rm_with_temperature,
             ncol = 2) 
```

XGBoost
Reformat dataset to match with XGBoost format
```{r}
# Without temperature
train_matrix_without_temperature <- xgb.DMatrix(data = 
  as.matrix(X_train_without_temperature), label = y_train$label)
test_matrix_without_temperature <- xgb.DMatrix(data = 
  as.matrix(X_test_without_temperature), label = y_test$label)

# With temperature
train_matrix_with_temperature <- xgb.DMatrix(data = 
  as.matrix(X_train_with_temperature), label = y_train$label)
test_matrix_with_temperature <- xgb.DMatrix(data = 
  as.matrix(X_test_with_temperature), label = y_test$label)
```

Train the XGBoost model 
```{r}
#The algorithm still can detect that there are only 2 classes
numberOfClasses <- 3 
xgb_params <- list("objective" = "multi:softprob",
                   "eval_metric" = "mlogloss",
                   "num_class" = numberOfClasses)

# Without temperature
xgb_model_without_temperature <- xgb.train(
  params = xgb_params, data = train_matrix_without_temperature, nrounds = 250)

# With temperature
xgb_model_with_temperature <- xgb.train(
  params = xgb_params, data = train_matrix_with_temperature, nrounds = 250)
```

Train performance
```{r}
# Without temperature
xgb_train_pred_without_temperature <- predict(
  xgb_model_without_temperature, newdata = train_matrix_without_temperature)

xgb_train_prediction_without_temperature <- matrix(
  xgb_train_pred_without_temperature, nrow = numberOfClasses, 
  ncol= length(xgb_train_pred_without_temperature) / numberOfClasses) %>%
  t() %>%
  data.frame() %>%
  mutate(label = y_train$label,
         max_prob = max.col(., "last") - 1)

confusionMatrix(
  data = factor(xgb_train_prediction_without_temperature$max_prob),
  reference = factor(xgb_train_prediction_without_temperature$label),
  mode = "everything")
```

```{r}
# With temperature
xgb_train_pred_with_temperature <- predict(
  xgb_model_with_temperature, newdata = train_matrix_with_temperature)

xgb_train_prediction_with_temperature <- matrix(
  xgb_train_pred_with_temperature, nrow = numberOfClasses, 
  ncol= length(xgb_train_pred_with_temperature) / numberOfClasses) %>%
  t() %>%
  data.frame() %>%
  mutate(label = y_train$label,
         max_prob = max.col(., "last") - 1)

confusionMatrix(data = factor(xgb_train_prediction_with_temperature$max_prob),
                reference = factor(xgb_train_prediction_with_temperature$label),
                mode = "everything")
```

Test performance
```{r}
# Without temperature
xgb_test_pred_without_temperature <- predict(
  xgb_model_without_temperature, newdata = test_matrix_without_temperature)

xgb_test_prediction_without_temperature <- matrix(
  xgb_test_pred_without_temperature, nrow = numberOfClasses, 
  ncol=length(xgb_test_pred_without_temperature)/numberOfClasses) %>%
  t() %>%
  data.frame() %>%
  mutate(label = y_test$label,
    max_prob = max.col(., "last") - 1)
```

```{r}
accuracy_xgb_without_temperature_pred_test <- mean(
  xgb_test_prediction_without_temperature$max_prob == 
    xgb_test_prediction_without_temperature$label)
accuracy_xgb_without_temperature_pred_test
```

```{r}
# Plot confusion matrix
cm_xgb_without_temperature_test <- as.data.frame(as.table(confusionMatrix(
  data = factor(xgb_test_prediction_without_temperature$max_prob),
  reference = factor(xgb_test_prediction_without_temperature$label),
  )))
cm_xgb_without_temperature_test <- cm_xgb_without_temperature_test %>% 
  rename(Actual = Reference)
cm_xgb_without_temperature_test$classes <- 2
cm_xgb_without_temperature_test$model <- "XGBoost"
cm_xgb_without_temperature_test$with_or_without_temperature <- 
  "Without Temperature"
cm_xgb_without_temperature_test$accuracy <- 
  accuracy_xgb_without_temperature_pred_test

# Remap labels
cm_xgb_without_temperature_test$Prediction <- as.numeric(
  cm_xgb_without_temperature_test$Prediction)
cm_xgb_without_temperature_test$Actual <- as.numeric(
  cm_xgb_without_temperature_test$Actual)
cm_xgb_without_temperature_test[
  cm_xgb_without_temperature_test$Prediction == 1,]$Prediction <- "Adversity"
cm_xgb_without_temperature_test[
  cm_xgb_without_temperature_test$Prediction == 2,]$Prediction <- "Affluent"
cm_xgb_without_temperature_test[
  cm_xgb_without_temperature_test$Actual == 1,]$Actual <- "Adversity"
cm_xgb_without_temperature_test[
  cm_xgb_without_temperature_test$Actual == 2,]$Actual <- "Affluent"

cm_xgb_without_temperature_test$Prediction <- factor(
  cm_xgb_without_temperature_test$Prediction, 
  levels = c("Adversity", "Affluent"))
cm_xgb_without_temperature_test$Actual <- factor(
  cm_xgb_without_temperature_test$Actual, 
  levels = c("Adversity", "Affluent"))

png("output/Confusion Matrix - XGBoost - 2 Classes - Without temperature.png",
    width = 125, height = 75, units='mm', res = 700)

ggplot(cm_xgb_without_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Accuracy: ", 
                      round(cm_xgb_without_temperature_test$accuracy, 3))) + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
confusionMatrix(
  data = factor(xgb_test_prediction_without_temperature$max_prob), 
  reference = factor(xgb_test_prediction_without_temperature$label),
  mode = "everything")
```

```{r}
# With temperature
xgb_test_pred_with_temperature <- predict(
  xgb_model_with_temperature, newdata = test_matrix_with_temperature)

xgb_test_prediction_with_temperature <- matrix(
  xgb_test_pred_with_temperature, nrow = numberOfClasses, 
  ncol=length(xgb_test_pred_with_temperature)/numberOfClasses) %>%
  t() %>%
  data.frame() %>%
  mutate(label = y_test$label,
    max_prob = max.col(., "last") - 1)
```

```{r}
accuracy_xgb_with_temperature_pred_test <- mean(
  xgb_test_prediction_with_temperature$max_prob == 
    xgb_test_prediction_with_temperature$label)
accuracy_xgb_with_temperature_pred_test
```

```{r}
# Plot confusion matrix
cm_xgb_with_temperature_test <- as.data.frame(as.table(confusionMatrix(
  data = factor(xgb_test_prediction_with_temperature$max_prob),
  reference = factor(xgb_test_prediction_with_temperature$label)
  )))
cm_xgb_with_temperature_test <- cm_xgb_with_temperature_test %>% 
  rename(Actual = Reference)
cm_xgb_with_temperature_test$classes <- 2
cm_xgb_with_temperature_test$model <- "XGBoost"
cm_xgb_with_temperature_test$with_or_without_temperature <- 
  "With Temperature"
cm_xgb_with_temperature_test$accuracy <- 
  accuracy_xgb_with_temperature_pred_test

# Remap labels
cm_xgb_with_temperature_test$Prediction <- as.numeric(
  cm_xgb_with_temperature_test$Prediction)
cm_xgb_with_temperature_test$Actual <- as.numeric(
  cm_xgb_with_temperature_test$Actual)
cm_xgb_with_temperature_test[
  cm_xgb_with_temperature_test$Prediction == 1,]$Prediction <- "Adversity"
cm_xgb_with_temperature_test[
  cm_xgb_with_temperature_test$Prediction == 2,]$Prediction <- "Affluent"
cm_xgb_with_temperature_test[
  cm_xgb_with_temperature_test$Actual == 1,]$Actual <- "Adversity"
cm_xgb_with_temperature_test[
  cm_xgb_with_temperature_test$Actual == 2,]$Actual <- "Affluent"

cm_xgb_with_temperature_test$Prediction <- factor(
  cm_xgb_with_temperature_test$Prediction, 
  levels = c("Adversity", "Affluent"))
cm_xgb_with_temperature_test$Actual <- factor(
  cm_xgb_with_temperature_test$Actual, 
  levels = c("Adversity", "Affluent"))

png("output/Confusion Matrix - XGBoost - 2 Classes - With temperature.png", 
    width = 125, height = 75, units='mm', res = 700)

ggplot(cm_xgb_with_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Accuracy: ", 
                      round(cm_xgb_with_temperature_test$accuracy, 3))) + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
confusionMatrix(
  data = factor(xgb_test_prediction_with_temperature$max_prob),
  reference = factor(xgb_test_prediction_with_temperature$label),
  mode = "everything")
```

```{r}
plot_xgb_without_temperature <- 
  ggplot(cm_xgb_without_temperature_test, 
         aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("Without temperature \n Accuracy: ", 
                      round(cm_xgb_without_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))

plot_xgb_with_temperature <- ggplot(cm_xgb_with_temperature_test, 
       aes(x = Actual, y = Prediction, fill = Freq)) + 
  geom_tile() + 
  coord_equal() +
  scale_fill_gradient() + 
  guides(fill = "none") + 
  geom_text(aes(label = Freq), color = "white") + 
  labs(title = paste0("With temperature \n Accuracy: ", 
                      round(cm_xgb_with_temperature_test$accuracy, 
                            3))) + 
  theme(plot.title = element_text(hjust = 0.5))

png("output/Confusion Matrix - XGBoost - 2 Classes - All.png",
    width = 175, height = 75, units='mm', res = 700)

grid.arrange(plot_xgb_without_temperature, 
             plot_xgb_with_temperature,
             ncol = 2) 
```

```{r}
# Compile confusion matrices
cm_all <- rbind(
  # Logistic regression
  cm_logistic_without_temperature_test, 
  cm_logistic_with_temperature_test,
  
  # # Decision Tree
  # cm_dtree_without_temperature_test, 
  # cm_dtree_with_temperature_test,
  
  # Random Forest
  cm_rf_without_temperature_test,
  cm_rf_with_temperature_test,
  
  # XGBoost
  cm_xgb_without_temperature_test,
  cm_xgb_with_temperature_test
  )
cm_all
```

```{r}
write.csv(cm_all, "output/cm_2.csv", row.names = FALSE)
```

